### 2022-01-20 plos_latex_template.tex
edited Rules 6–7 ~ @jananiravi

% Be more direct
Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. Including people with disabilities in the organizing team can have a large impact from the beginning (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making \cite{irishIncreasingParticipationUsing2020}. As other aspects of inclusion, dealing with accessibility at the last minute is the recipe for a disastrous conference. If you do not consider accessibility from the conference inception, it is highly likely that you will be better off without trying to patch accessibility at the last minute. Key decisions in this respect are hard to correct.

If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live interpretation, just to name a few important features. In addition to this, the organizers should take care proactively of invisible disabilities (e.g. dyslexia, anxiety, ADHD, autism). This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.

Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly, keyboard accessible, and have alternative text. Every video should have good-quality (not automated) captions and a transcript; live presentations should have good-quality captions too. 
Provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have. For instance, useR! 2021 published accessibility guidelines for every format of the conference, a blog post with recommendations, and two post-conference blog posts regarding the whole accessibility experience. The participation of a blind person in the organizing team was key to several of these decisions. 
Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.

@@

Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection, etc. Accessibility efforts should be explicitly displayed in the official communication channels and inform potential attendees with disabilities a point of contact to solve any questions in this respect.
Clear communication about the lack of accessibility of some spaces or activities is also highly recommended, to give a clear notion of the conference experience (see \textbf{Rule 8} about communication strategy).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international meetings, the linguistic diversity of the participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpages, and communications. 
While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.%RJ: Could cite Amano's paper here.
Non-native English speakers could miss opportunities to attend or actively participate in conferences (e.g. asking questions or participating in discussions),
and conferences may in turn miss innovative contributions.

% Rocío: main idea: make conferences more linguistically inclusive
Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (\textbf{Rule 2} about a welcoming environment for all). This may be done at different levels.
Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language. That way, the abstract will be judged primarily by the quality or relevance of the work instead of the English grammar.
For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities) follow the presentations. Whenever possible, identify other key languages for the conference and provide translated captions or live interpretation into these key languages, including multilingual Q \& A sessions.
The following step would be having sessions and events in languages other than English, both regarding the technical content of the conference and the social and networking aspects. Networking sessions in languages other than English can include captions, interpretation or none of them; the latter has the benefit of not requiring additional technical adjustments. 
Keynotes and regular talks should always be accompanied by captions or live interpretation to English.
Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are available. 
Remind your native English speaking audience to be respectful of accents and mistakes in English.

>>>>

% Be more direct
Unfortunately, even today, conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. Including people with disabilities in the organizing team can have a large impact from the beginning (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making \cite{irishIncreasingParticipationUsing2020}. As with other aspects of inclusion, dealing with accessibility at the last minute is unlikely to yield much success. There are several faces of accessibility that organizers must address and accommodate. Therefore, accessibility must be considered right from the conference inception phase, since it's difficult to course correct at later stages.

If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and live translation, to name a few. In addition to this, the organizers should proactively take care of invisible disabilities (e.g. dyslexia, anxiety, ADHD, autism). This could be done in many ways, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.

Regardless of the conference format, all platforms (website, chat, conference administration tools) and images used for the communication strategy of the conference should be screen-reader friendly, keyboard accessible, and have alternative text. Every video should have good-quality (not automated) captions and a transcript; live presentations should have good-quality captions too. 
Provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions presenters and attendees may have. For instance, useR! 2021 published accessibility guidelines for every format of the conference, a blog post with recommendations, and two post-conference blog posts regarding the whole accessibility experience. The participation of a blind person in the organizing team was key to the successful implementation of several of these decisions. 
Slide decks should be made available beforehand, either in webpages or available for download to ensure that everybody can follow their content during the presentations.

@@

Importantly, all these accessibility practices are inclusive not only for people with disabilities but to everyone.
For instance, captions are helpful for non-native speakers, having the material available for download helps attendees with low bandwidth connection. Accessibility efforts should be explicitly displayed in the official communication channels and potential attendees with disabilities must be provided with a point of contact to help clarify any questions.
Clear communication about the lack of accessibility for some spaces or activities is also highly recommended, to give a clear notion of the conference experience and accommodation (see \textbf{Rule 8} about communication strategy).

@@

% Rocío: main idea: English as the only language makes some people privileged and is a barrier for others
In international meetings, the linguistic diversity of the participants is often overlooked. 
English is usually the official and sole language for submissions, presentations, tutorials, workshops, conference platforms, webpages, and communications. 
While English is indeed regarded as the primary language in scientific communication and one official language makes it conducive to communicate widely, this makes being a native English speaker a privilege.%RJ: Could cite Amano's paper here.
Non-native English speakers may miss opportunities to attend or actively participate in conferences (e.g. asking questions or partaking in discussions),
and conferences may in turn miss innovative contributions.

% Rocío: main idea: make conferences more linguistically inclusive
Providing a welcoming and diverse environment by encouraging the full participation of non-native English speakers is critical (\textbf{Rule 2} about a welcoming environment for all). This may be done at different levels.
\emoji{bulb} Allow for abstract submission in both English and the language the person feels more comfortable with, and consider the possibility of assigning a reviewer who is fluent in that language. That way, the abstract will be judged primarily by the quality or relevance of the work instead of English proficiency.
For presentations spoken in English, provide English-to-English captions to help non-native English speakers (in addition to people with hearing disabilities) follow the presentations. Whenever possible, identify other prominent languages for the conference and provide translated captions or live interpretation into these key languages, including multilingual Q \& A sessions.
Another way to engage non-native English participants is to host sessions and events in languages other than English, both technical and social/networking aspects. Networking sessions in other languages can include captions, interpretation or none of them; the latter has the benefit of not requiring additional technical adjustments. 
Keynotes and regular talks should always be accompanied by captions or live translation to English.
Promote the attendance to presentations in languages other than English as key components of your schedule and announce when captions are made available. 
To ensure inclusion without language barriers, remind your native English-speaking audience to be respectful of accents and possible grammatical mistakes, and embrace the backgrounds of non-English speakers while focusing on the core content.
